On Etiology of Cybersecurity

### 2020-06-08 reports/conference_paper/main.tex
first full draft of the conference paper

\documentclass[conference]{IEEEtran}
\usepackage[table]{xcolor}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[style=numeric]{biblatex}
\usepackage{amsmath}
\addbibresource{../literature.bib}% Syntax for version >= 1.2
\usepackage{graphicx}
\graphicspath{{../images/}}
\usepackage{multirow}

@@

\newcommand{\assertionRegion}{\mathcal{A}}
\newcommand{\beliefRegion}{\mathcal{B}}
\newcommand{\factRegion}{\mathcal{F}}
\newcommand{\rcc}{rcc}
\newcommand{\agentuniverse}{Ag}
\newcommand{\abf}{\assertionRegion,\beliefRegion,\factRegion}
\newcommand{\Rcc}[2]{rcc(#1,#2)}
\newcommand{\connects}[2]{C(#1,#2)}
\newcommand{\disconnected}[2]{\neg C(#1,#2)}
\newcommand{\partof}[2]{P(#1,#2)}
\newcommand{\overlaps}[2]{O(#1,#2)}
\newcommand{\eq}[2]{EQ(#1,#2)}
\newcommand{\pp}[2]{PP(#1,#2)}
\newcommand{\po}[2]{PO(#1,#2)}
\newcommand{\ppi}[2]{PPi(#1,#2)}
\newcommand{\dr}[2]{DR(#1,#2)}
\newcommand{\all}[2]{T(#1,#2)}
\newcommand{\rassert}[3]{\mathcal{A}_{#1\rightarrow #2}#3}
\newcommand{\lassert}[3]{\mathcal{A}_{#1\leftarrow #2}#3}

\newtheorem{definition}{Definition}%[section]

@@

\begin{abstract}
The abstract goes here.
\end{abstract}

@@

\subsection{Systems and Multi-Agent Systems}\label{sec:system}
The ISO/IEC/IEEE 15288:2015 (System Life Cycle Processes) provides a definition
of system as ``A combination of interacting elements organized to achieve one
or more stated purposes.''\autocite{ISO201515288}.  Therefore, a system can be
considered as a single agent where its interacting elements are the
constituents of the agent itself. For the sake of simplicity we first define a
system as an agent and then extend the definition to a ``combination of
interacting'' agents.

@@

\subsection{Qualitative Evaluation of Agent Space in $\abf$}\label{sec:agentspace}
While a quantitative analysis reveals how many possible configurations of an
agent (i.e. a system) exist w.r.t. the $\abf$ theory (i.e. $54/125$ in RCC5), a
qualitative analysis of the different configurations describes the
configurations allowed by the $\abf$ theory, and how those configurations can
be categorized.  In Table~\ref{tab:5com}, we provide the generic composition
table of RCC5 over 3 regions instantiated over $\abf$, which shows the whole
state space for a single agent. The color coding of the table represents the
security risk related to a generic agent.

@@

\section{Prediction of Security Weaknesses}\label{sec:theory}
\begin{figure*}[t]
	\centering
	\includegraphics[width=.9\textwidth]{mashyp.pdf}
	\caption{asd}
	\label{fig:mashyp}
\end{figure*}

\paragraph{Fun, Behavior, Channels}

\section{Security Risk Assessment}\label{sec:secra}
\paragraph{Automated Security Requirements Extraction}

\section{Discussions}\label{sec:discussion}
\paragraph{Framework for Theory Falsification}
\paragraph{SPDL, Testing, and Falsification}
\paragraph{Relation with DY}
\paragraph{Abstraction Level}

\section{Conclusion and Future Works}
\begin{itemize}
	\item Verification
	\item Testing
\end{itemize}

\section*{Acknowledgment}
We thank Katia Santac\`a for the development of the initial ideas, and for
the helpful discussions that allowed us to make this step forward.

>>>>

\documentclass[conference]{IEEEtran}
\usepackage[table]{xcolor}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[style=numeric]{biblatex}
\usepackage{amsmath}
\addbibresource{../literature.bib}% Syntax for version >= 1.2
\usepackage{graphicx}
\graphicspath{{../images/}}
\usepackage{multirow}
\usepackage{enumitem}
\usepackage{tikz}
\usepackage{amsthm}
\usepackage{rotating}

@@

\newcommand{\assertionRegion}{\mathcal{A}}
\newcommand{\beliefRegion}{\mathcal{B}}
\newcommand{\factRegion}{\mathcal{F}}
\newcommand{\rcc}{rcc}
\newcommand{\agentuniverse}{Ag}
\newcommand{\abf}{\assertionRegion,\beliefRegion,\factRegion}
\newcommand{\abftheory}{\assertionRegion\beliefRegion\factRegion}
\newcommand{\Rcc}[2]{rcc(#1,#2)}
\newcommand{\connects}[2]{C(#1,#2)}
\newcommand{\disconnected}[2]{\neg C(#1,#2)}
\newcommand{\partof}[2]{P(#1,#2)}
\newcommand{\overlaps}[2]{O(#1,#2)}
\newcommand{\eq}[2]{EQ(#1,#2)}
\newcommand{\pp}[2]{PP(#1,#2)}
\newcommand{\po}[2]{PO(#1,#2)}
\newcommand{\ppi}[2]{PPi(#1,#2)}
\newcommand{\dr}[2]{DR(#1,#2)}
\newcommand{\all}[2]{T(#1,#2)}
\newcommand{\rassert}[3]{\mathcal{A}_{#1\rightarrow #2}#3}
\newcommand{\lassert}[3]{\mathcal{A}_{#1\leftarrow #2}#3}

\newcommand*\halfbullet[1][0.4ex]{%
  \begin{tikzpicture}
	  \raisebox{0.6pt}{\draw[fill] (0,0)-- (90:#1) arc (90:270:#1) -- cycle ;
	  \draw (0,0) circle (#1);}
  \end{tikzpicture}}

\newcommand{\fakedmapsfromchar}{%
  \mathrel{\mathpalette\reflectmathchar\mapstochar}%
}
\makeatletter
\newcommand{\reflectmathchar}[2]{%
  \reflectbox{$\m@th#1#2$}%
}
\makeatother
\newcommand{\rightbararrow}{\rightarrow\fakedmapsfromchar}
\newcommand{\leftbararrow}{\mapstochar\rightarrow}
\newcommand{\rightleftbararrow}{\mapstochar\rightarrow\fakedmapsfromchar}
\setlength{\fboxrule}{0.1pt}
\setlength{\fboxsep}{1.5pt}
\newcommand{\secureportelem}{\mathrel{\rightarrow}}
\newcommand{\secureport}{\fbox{\ensuremath \secureportelem}}
\newcommand{\dropportelem}{\mathrel{\circ\!\!\rightarrow}}
\newcommand{\dropport}{\fbox{\ensuremath \dropportelem}}
\newcommand{\insertportelem}{\mathrel{\bullet\!\!\rightarrow}}
\newcommand{\insertport}{\fbox{\ensuremath \insertportelem}}
\newcommand{\injectionportelem}{\mathrel{\halfbullet\!\!\rightarrow}}
\newcommand{\injectionport}{\fbox{\ensuremath \injectionportelem}}
\newcommand{\selectiveportelem}{\mathrel{\leftbararrow}}
\newcommand{\selectiveport}{\fbox{\ensuremath \selectiveportelem}}
\newcommand{\selectiveinsertportelem}{\mathrel{\bullet\!\!\leftbararrow}}
\newcommand{\selectiveinsertport}{\fbox{\ensuremath \selectiveinsertportelem}}
\newcommand{\selectiveinjectionportelem}{\mathrel{\halfbullet\!\leftbararrow}}
\newcommand{\selectiveinjectionport}{\fbox{\ensuremath \selectiveinjectionportelem}}
\newcommand{\selectivedropportelem}{\mathrel{\circ\!\!\leftbararrow}}
\newcommand{\selectivedropport}{\fbox{\ensuremath \selectivedropportelem}}
%communication channels
%secure input port
\newcommand{\sdcom}{\mathrel{\rightarrow\!\!\circ}}
\newcommand{\sicom}{\mathrel{\rightarrow\!\!\bullet}}
\newcommand{\sjcom}{\mathrel{\rightarrow\!\!\halfbullet}}
\newcommand{\ssdcom}{\mathrel{\rightbararrow\!\!\circ}}
\newcommand{\ssicom}{\mathrel{\rightbararrow\!\!\bullet}}
\newcommand{\ssjcom}{\mathrel{\rightbararrow\!\halfbullet}}
%drop input port
\newcommand{\ddcom}{\mathrel{\circ\!\!\rightarrow\!\!\circ}}
\newcommand{\dicom}{\mathrel{\circ\!\!\rightarrow\!\!\bullet}}
\newcommand{\dscom}{\mathrel{\circ\!\!\rightarrow}}
%insert input port
\newcommand{\idcom}{\mathrel{\bullet\!\!\rightarrow\!\!\circ}}
\newcommand{\iicom}{\mathrel{\bullet\!\!\rightarrow\!\!\bullet}}
\newcommand{\ijcom}{\mathrel{\bullet\!\!\rightarrow\!\!\halfbullet}}
\newcommand{\isdcom}{\mathrel{\bullet\!\!\rightbararrow\!\!\circ}}
\newcommand{\isicom}{\mathrel{\bullet\!\!\rightbararrow\!\!\bullet}}
\newcommand{\isjcom}{\mathrel{\bullet\!\!\rightbararrow\!\halfbullet}}
\newcommand{\iscom}{\mathrel{\bullet\!\!\rightarrow}}
%injection input
\newcommand{\jdcom}{\mathrel{\halfbullet\!\!\rightarrow\!\!\circ}}
\newcommand{\jicom}{\mathrel{\halfbullet\!\!\rightarrow\!\!\bullet}}
\newcommand{\jjcom}{\mathrel{\halfbullet\!\!\rightarrow\!\!\halfbullet}}
\newcommand{\jsdcom}{\mathrel{\halfbullet\!\!\rightbararrow\!\!\circ}}
\newcommand{\jsicom}{\mathrel{\halfbullet\!\!\rightbararrow\!\!\bullet}}
\newcommand{\jsjcom}{\mathrel{\halfbullet\!\!\rightbararrow\!\halfbullet}}
\newcommand{\jscom}{\mathrel{\halfbullet\!\!\rightarrow}}
%selective drop input
\newcommand{\sddcom}{\mathrel{\circ\!\!\leftbararrow\!\!\circ}}
\newcommand{\sdicom}{\mathrel{\circ\!\!\leftbararrow\!\!\bullet}}
\newcommand{\sdjcom}{\mathrel{\circ\!\!\leftbararrow\!\!\halfbullet}}
\newcommand{\sdsdcom}{\mathrel{\circ\!\!\rightleftbararrow\!\!\circ}}
\newcommand{\sdsicom}{\mathrel{\circ\!\!\rightleftbararrow\!\!\bullet}}
\newcommand{\sdsjcom}{\mathrel{\circ\!\!\rightleftbararrow\!\halfbullet}}
\newcommand{\sdscom}{\mathrel{\circ\!\!\leftbararrow}}
%selective insertion
\newcommand{\sidcom}{\mathrel{\bullet\!\!\leftbararrow\!\!\circ}}
\newcommand{\siicom}{\mathrel{\bullet\!\!\leftbararrow\!\!\bullet}}
\newcommand{\sijcom}{\mathrel{\bullet\!\!\leftbararrow\!\!\halfbullet}}
\newcommand{\sisdcom}{\mathrel{\bullet\!\!\rightleftbararrow\!\!\circ}}
\newcommand{\sisicom}{\mathrel{\bullet\!\!\rightleftbararrow\!\!\bullet}}
\newcommand{\sisjcom}{\mathrel{\bullet\!\!\rightleftbararrow\!\halfbullet}}
\newcommand{\siscom}{\mathrel{\bullet\!\!\leftbararrow}}
%selective injection
\newcommand{\sjdcom}{\mathrel{\halfbullet\!\!\leftbararrow\!\!\circ}}
\newcommand{\sjicom}{\mathrel{\halfbullet\!\!\leftbararrow\!\!\bullet}}
\newcommand{\sjjcom}{\mathrel{\halfbullet\!\!\leftbararrow\!\!\halfbullet}}
\newcommand{\sjsdcom}{\mathrel{\halfbullet\!\!\rightleftbararrow\!\!\circ}}
\newcommand{\sjsicom}{\mathrel{\halfbullet\!\!\rightleftbararrow\!\!\bullet}}
\newcommand{\sjsjcom}{\mathrel{\halfbullet\!\!\rightleftbararrow\!\halfbullet}}
\newcommand{\sjscom}{\mathrel{\halfbullet\!\!\leftbararrow}}

\newtheorem{definition}{Definition}%[section]
\newtheorem{hypothesis}{Hypothesis}%[section]

@@

\begin{abstract}
The objective of this research is the development of a theory that defines (all
	and only) the possible insecurity and security configurations of any
	abstract system. The theory is structured upon other theories that
	defines how a component of a system can be abstracted into an agent,
	defining how agents can be formalized (both syntactically and
	semantically) to describe an abstract system, such as a graph. The core
	theories are the epistemological definition of knowledge, Beliefs, and
	Information, the Assertion-Belief-Fact framework, and mereotopology.
	We implemented a formal theory (of axioms) of a mereotopology, and of
	the Region Connection Calculus (RCC3 and RCC5) in a Python program that
	uses the Z3 SMT solver. The results show that a single component (i.e.
	agent) of an abstract system has a definite number of  different
	insecurity configurations (e.g. 53 using RCC5 over a topological
	structure) and only 1 secure (i.e.  expected) configuration. The
	configurations are reported as models satisfying the abstract system
	semantics. We use these results to predict all possible security
	weaknesses of a system, instead of relying on databases of known
	weaknesses, and to lay the foundations of a scientific theory of
	cybersecurity.  We show a concrete applications of our theory to the
	risk assessment of an ad-hoc system.
\end{abstract}

@@

\subsection{Systems as Multi-Agent Systems}\label{sec:system}
The ISO/IEC/IEEE 15288:2015 (System Life Cycle Processes) provides a definition
of system as ``A combination of interacting elements organized to achieve one
or more stated purposes.''\autocite{ISO201515288}.  Therefore, a system can be
considered as a single agent where its interacting elements are the
constituents of the agent itself. For the sake of simplicity we first define a
system as an agent and then extend the definition to a ``combination of
interacting'' agents.

@@

\subsection{Qualitative Evaluation of Agent Space in $\abf$}\label{sec:agentspace}

@@

While a quantitative analysis reveals how many possible configurations of an
agent (i.e. a system) exist w.r.t. the $\abf$ theory (i.e. $54/125$ in RCC5), a
qualitative analysis of the different configurations describes the
configurations allowed by the $\abf$ theory, and how those configurations can
be categorized.  In Table~\ref{tab:5com}, we provide the generic composition
table of RCC5 over 3 regions instantiated over $\abf$, which shows the whole
state space for a single agent. The color coding of the table represents the
security risk related to a generic agent.

\section{Prediction of Security Weaknesses}\label{sec:theory}
Several standards mandates a secure-by-design approach in which cybersecurity
shall be considered at the very early stages of the design process.  For
example,
\begin{enumerate}[noitemsep]
	\item DO-326A -- ``Airworthiness Security Process Specification''
		requires a cybersecurity risk assessment of the design and is
		the ``are the only Acceptable Means of Compliance (AMC) by FAA
		\& EASA for aviation cyber-security airworthiness certification,
		as of 2019'' as pointed out by SAE in \autocite{SAE2019DO326A}.
	\item NIST 800-82 \autocite{Stouffer2011guide} -- ``guide to Industrial
		Control System (ICS) Security''
	\item J3061:2016-1 \autocite{SAE2016J3061} -- ``Cybersecurity Guidebook
		for Cyber-Physical Vehicle Systems'' defines ``set of
		high-level guiding principles for Cybersecurity as it relates
		to cyber-physical vehicle systems'' and states that
		``incorporate Cybersecurity into cyber-physical vehicle systems
		from concept phase through production, operation, service, and
		decommissioning''
\end{enumerate}
However, standards do not describe in detail how to perform a security risk 
assessment and only vaguely define the overall objective. The overall
objective of this first step of the cybersecurity engineering process
is to provide an understanding of the potential security risks.
Roughly speaking, a security risk assessment (a.g. \autocite{coras,secram}) 
methodology starts from the (i) identification of assets, (ii) determines
the threat scenarios (correlating vulnerabilities, threats and incidents) e.g. 
threat diagrams in the CORAS approach, (iii) calculate the risk as a function
between impact and likelihood of threat scenarios, and (iv) finally provides
treatments (mandating a partial re-design) and residual risks.
All the methodologies and tools we reviewed (Threatmodeler, CORAS, SECRAM, ...)
relies on the expertise of the person who performs the risk assessment for
the identification of threats and for the quantitative estimation of risks.
As Herley puts it \autocite{cite}, ``we don't have a test for security and
we don't have a metric for security'' while standards requires a security risk assessment
which, in turn, mandates a metric to evaluate the security risk.

So far, we have reasoned on systems in the abstract, considering
generic MAS; mapping Epistemological concepts
to MAS. We now map MAS to CPS and show how this mapping allows us
to predict security weaknesses in a system.

\subsection{From Multi-Agent to Cyber-Physical Systems}
\begin{figure*}[t]
	\centering
	\includegraphics[width=.9\textwidth]{mashyp.pdf}
	\caption{asd}
	\label{fig:mashyp}
\end{figure*}

As depicted in Figure~\ref{fig:mashyp}, we relate MAS and CPS as follows.
\begin{itemize}
	\item We consider a System as an hierarchy of agents. So, we map agents
		to \emph{systems, sub-systems, or devices}, depending on the
		granularity of the design. For example, a modeler can model a
		specific device, as in Figure~\ref{fig:mashyp}. In this case,
		the system is not decomposed into sub-systems or devices and
		the device is considered as an agent.
	\item Agents reason over Beliefs (i.e. transform Beliefs into other
		Beliefs) and for each \emph{component of a CPS} (system, sub-system,
		or device) is composed by a \emph{functional architecture} that
		transforms input-Beliefs into output-Beliefs.
	\item Components of a CPS have \emph{ports} to exchange information
		with the outer environment (which may be a sub-system),
		similarly to Agents. The transfer of information in a CPS is
		defined by \emph{channels}.
	\item The concept of Knowledge is related to the \emph{requirements}
		that, in a factorial way, describe how the CPS shall behave and
		its \emph{physical architecture}.

\paragraph{Input and Output Ports}
\begin{figure}[t]
	\centering
	\includegraphics[width=\columnwidth]{IOports.pdf}
	\caption{Input and Output Port of an agent}
	\label{fig:ioports}
\end{figure}
Since the $\abf$-theory is a theory of agents, we could consider ports as
agents that allows the exchange of information between a channel and another
agent.  However, we considered a port as a special type of agent that to avoid
an \emph{infinite regress} in which a port needs a port to transfer information
between the outside of the port to the inside of itself.

\begin{definition}{\bf Input or Output Port --}\label{def:port} 
	A port forwards information from the outside of an agent's boundary to
	the inside (input-port) or vice versa (output-port).  There exist two
	types of ports with the following behavior: an input-port changes
	transforms Assertions from a sender $s$ ($\rassert{s}{a}$) to Beliefs
	of an agent $a$ ($\beliefRegion_a$), while an output-port from belief
	to assertion.
\end{definition}
The \emph{quality of a port} is determined by the $\rcc$ relation between the
assertions received or sent and the belief, i.e.
$\Rcc{\assertionRegion_s}{\beliefRegion_a}$ for the input-port or
$\Rcc{\assertionRegion_r}{\beliefRegion_a}$. A port is, in fact, syntactic sugar to express the 
relation between Assertions and Beliefs. We only define secure input port
but the definition of a secure output port is symmetrical.

\begin{definition}{\bf Secure Input Port --}\label{def:secport}
	A secure input port ($s\secureport a$) always allow information as
	incoming assertions to flow from a sender $s$ to the behavior of the
	recipient $a$ (agent).  
	%If $p_I$ is a secure input port,
	%$\world\models\varphi$, $\varphi\in\rassert{s}{a}$, and $a$ has the
	%only input port $p_I$, $\world'\models\beliefRegion_a$ and
	%$\varphi\in\beliefRegion_a$ for any $\world\modalrelation\world'$.
\end{definition}

\paragraph{Port Weaknesses}
From our definition of ports, it follows that there exist only the following \emph{six}
types of weaknesses, generating six types of insecure port in RCC5 (the
notation is reported for input-ports, i.e. on the lef-hand side of the arrow):
\begin{enumerate}[start=1, label={W\arabic*)}]
	\item \emph{replace port} ($s\dropport a$), where assertions reaches
		the port but do not pass the boundary of the agent (i.e. do not
		become belief of the agent)
	\item \emph{drop port} ($s\dropport a$), where assertions reaches the
		port but do not pass the boundary of the agent (i.e. do not
		become belief of the agent)
	\item \emph{insertion port} ($s\insertport a$), where some new
		information is believed by $a$ as incoming from the port but
		$s$ didn't send it
	\item \emph{injection port} ($s\injectionport a$), where information
		coming from $s$ is substituted with new information which
		becomes believed by $a$
	\item \emph{selective port}, where some information passes the port and
		part is either:
	\begin{enumerate}[start=1, label={W4.\arabic*)}]
		\item drop ($s\selectivedropport a$), 
		\item drop+insert ($s\selectiveinsertport a$).
	\end{enumerate}
\end{enumerate}

\fixnote{mr}{this proof needs to be restructured as in slides v-research-demo}
\begin{proof}
An input port is, in the $\abf$-theory, defined secure as long as the relation
	between the two regions of input assertions $\assertionRegion$ and
	output beliefs $\beliefRegion$ are equal, i.e.
	$\eq{\assertionRegion}{\beliefRegion}$. Therefore, any other relation
	should result in a weakness (related to an insecurity flaw) of that
	input port.  Using RCC5, there exist exactly other $4$ different type
	of relations, one of which is the discrete-from (DR) relation, i.e.
	$\dr{\assertionRegion}{\beliefRegion}$. When two regions are related by
	the DR relations, they have no subregion in common. Lets
	define a function weight $|X|$ such as, for any region $X$, it
	represents the smallest possible cardinality of a (mereo)topological base for
	$X$; where a base is a collection of regions in a (mereo)topology such that
	every open region can be written as union of elements of that base. 
	We distinguish between Regions that contain Information and Regions that
	don't by writing the latter as $\emptyset$.
	\begin{enumerate}
		\item if $\eq{\assertionRegion}{\beliefRegion})$ then either
			$\assertionRegion=\beliefRegion=\emptyset$ (no
			communication) or
			$\assertionRegion=\beliefRegion\neq\emptyset$ (forward
			communication)
		\item if $\dr{\assertionRegion}{\beliefRegion})$ then
			$\assertionRegion=\emptyset\oplus\beliefRegion=\emptyset$
			(we call \emph{insert} the former and \emph{full drop}
			the latter case), or
			$\assertionRegion\neq\emptyset\wedge\beliefRegion\neq\emptyset\wedge\assertionRegion\neq\beliefRegion$
			which we call \emph{replace} (i.e. drop and insert)
		\item if $\pp{\assertionRegion}{\beliefRegion})$ then
			$\beliefRegion$ contains and extend $\assertionRegion$
			which we call \emph{injection}
		\item if $\ppi{\assertionRegion}{\beliefRegion})$ then
			$\assertionRegion$ contains and extend $\beliefRegion$
			which we call \emph{selective drop}
		\item if $\po{\assertionRegion}{\beliefRegion})$ then a part of
			the $\assertionRegion$ is contained in the
			$\beliefRegion$ which is a combination of
			\emph{selective drop and generation}
	\end{enumerate}
\end{proof}

\paragraph{Communication Channels}
\begin{figure}[t]
	\centering
	\includegraphics[width=\columnwidth]{channel.pdf}
	\caption{Communication over a Mono-directional Channel}
	\label{fig:channel}
\end{figure}
In this work, we only consider
mono-directional channels and communication but the extension to bi-directional
channel can be considered as the union of two unidirectional channels. A
mono-directional channel is defined by the assertions sent or received (over
the channel).

We start by considering the difference between a (communication)
mono-directional channel (channel from now on) and an agent, as we did for the
ports, since the $\abf$-theory is a theory of agents.  In fact, if a channel
were considered an agent (channel-agent) then the question would be how an
agent would transfer its Assertions to the channel-agent. If the channel
between the agent and the channel-agent is again an agent, we would generate an
\emph{infinite regress}. Therefore, we do allow channel-agents but we assume a
finite depth (of detail) for a channel, where there exists a bottom-channel
which is not an agent. For now, we do not constrain a channel-agent in any way
so there is no difference between a channel agent and agent. Therefore,
we consider channels to be bottom-channels,
defined as agents with the pre-defined behavior (i.e. defined in a axiomatic
way) of forwarding any input-assertion as output-assertion, without modifying it\footnote{Nothing
prevents us from introducing additional constraints to the channel as storing
Assertions that are transferred over the channel, or filter out some
input-assertions.}.

\begin{definition}{\bf Secure Mono-directional Channel (bottom-channel) --}\label{def:monochannel}
	A mono-directional channel between the two agents ($s \rightarrow r$) is defined as 
	an agent which behavior is (dogmatically) defined as: to forward any Assertion received from $s$ over an input-port, to  
	the output-port where $r$ is listening to.
\end{definition}
The \emph{quality of a mono-directional} channel is defined as the $\rcc$
relation between the Assertions made by the sender and the ones received by the
receiver, i.e. $\Rcc{\assertionRegion_s}{\assertionRegion_r}$.

\paragraph{Channel Weaknesses}
Given that a mono-directional bottom-channel is assumed to be perfectly
forwarding any Assertion from its input-port to its output-port, there is no
insecure behavior but only the combination of the weaknesses of the input and
output port; therefore there exists $(7^2)-2=47$ theoretical configurations
($7^2$ because there are $6$ insecure types of port, plus $1$ secure type, on
both input and output side; and we exclude the configuration with $2$ secure
types as input and output, $-2$); where only 43 are possible. For the sake of
readability, we report 6 examples in the following but the proof by exhaustion
over all the possible cases is straightforward.

\begin{enumerate}[start=6, label={W\arabic*)}]
	\item \emph{secure output port} and \emph{input drop port} ($s\sdcom r$), 
	\item \emph{secure output port} and \emph{input insertion port} ($s\sicom r$), 
	\item \emph{output drop port} and \emph{input drop port} ($s\ddcom r$)
	\item \emph{output drop port} and \emph{input insertion port} ($s\dicom r$)
	\item \emph{output drop port} and \emph{input secure port} ($s\dscom r$)
	\item \emph{output injection port} and \emph{input secure port} ($s\jscom r$)
\end{enumerate}

\paragraph{Security Weaknesses} 
It is important to note that all the results of the application of the
$\abf$-theory to channels (the analysis of the relation
$\Rcc{\assertionRegion_o}{\assertionRegion_i}$) lead to the same results of the
analysis of a pair of an input and output port (i.e.
$\Rcc{\beliefRegion_1}{\assertionRegion_2}$ and
$\Rcc{\assertionRegion_2}{\beliefRegion_3}$).  The same result can be obtained
by analyzing the relation between the outputs of a functional block and the
inputs of another functional block, where functional blocks are constituents of
the functional architecture as described afterwards.  As depicted in
Figure~\ref{fig:atom}, so far, we have considered Information generated by a
Port $P_I$ and then sent through a channel $C$ to another (recipient) port
$P_O$. In this scenario where Ports and then Channels are atomic (otherwise
raising infinite regress), we can only  consider the relations between Ports
and Channel; considering both input-port to channel and channel to output-port.
In fact, the Weaknesses of a channel are defined in terms of the Weaknesses of
Ports. We now introduce the concept of functional architecture and its
decomposition into functional blocks.  Io order to define a functional block
without encountering a infinitely recursive definition, we must reach the same
conclusions as for the Channel. Therefore, describing the information as
flowing over a channel or in a functional block is purely syntactic sugar.

We can summarize these results by saying that the relations between Assertions
and Beliefs, Output Assertions of an Agent and Input Assertions of another
Agent, or Output Beliefs of a Functional Block and Input Beliefs of another
Block can \emph{only} be affected by the following weaknesses: replace, drop,
injection, insertion, selective drop, and selective drop + insertion.  We can,
then, deduce the following security properties to mitigate security weaknesses
of a port or a channel (between ports, functional blocks, or both).
\begin{itemize}
	\item \emph{Order-preserving} -- it shall be known if Information is \emph{replaced}
	\item \emph{Availability} -- it shall be known if Information is \emph{dropped} or \emph{selectively dropped}
	\item \emph{Integrity} -- it shall be known if Information is \emph{injected}
	\item \emph{Authentication} -- it shall be known if Information is \emph{inserted}
\end{itemize}
%We can also generalize the weaknesses related to the data flow (i.e. the flow
%of Information as Assertions or Beliefs) as the relation between output-input
%Beliefs between functional blocks, output-input Assertions of a channel,
%inputs-outputs of a port, as follows.
%\begin{itemize}
%	\item \emph{equal}, expected nominal behavior
%	\item \emph{discrete-from}, drops all the inputs and inserts new malicious data
%	\item \emph{proper part}, selectively drops inputs
%	\item \emph{proper part inverse}, forwards all the inputs but crafts and
%	      inserts new malicious data
%	\item \emph{partial overlap}, selectively drops inputs and inserts new data.
%\end{itemize}

\begin{figure}[t]
	\centering
	\includegraphics[width=.9\columnwidth]{engineering_relations.pdf}
	\caption{Relations between Ports and Channel}
	\label{fig:atom}
\end{figure}

\paragraph{Functional Architecture}
A Functional Architecture takes Information as input-Beliefs and transforms the
Information into output-Beliefs. Those transformations occurs within the
Functional Architecture, where Functional Blocks transforms Beliefs into other
Beliefs. Similarly to channels, we could consider a Functional Block as a
Functional Architecture occurring in an infinite regress. Therefore, we
consider Functional Blocks as executing an abstract undefined behavior, of
which we only observe the inputs and the resulting outputs (Beliefs).

\begin{definition}{\bf Functional Block and Architecture --}\label{def:funblock}
	A functional block of an agent 
	takes a region of input-Beliefs and outputs a region of output-Beliefs. 
	A functional architecture is an interconnected system of functional blocks.
\end{definition}
It is evident that the quality of a functional block cannot be determined
by the difference between its inputs and outputs (as we did for
ports and channels). The behavior of a functional block
cannot be determined in general; since any functional block will have 
its own purpose based on functional requirements. 
Therefore, while a port semantics is determined by the relation 
between Assertions and Beliefs, the semantics of a Functional Block 
is determined by the relation between Facts/Requirements and I/O Beliefs.

In other words, a functional block is a generic agent with no pre-defined general
behavior (while ports and channels have a pre-defined behavior).

\begin{itemize}
	\item $\po{\beliefRegion}{\factRegion}$ the component has a Byzantine behavior where occasionally outputs the expected
	        output given the correct inputs. Not all the inputs are handled
	        properly, nor all the expected outputs are always generated
	        when correct inputs are given.
	\item $\pp{\beliefRegion}{\factRegion}$ part of the expected outputs are not
	        generated in response to the correct
	        inputs
	\item $\ppi{\beliefRegion}{\factRegion}$ the components
	        correctly performs the expected behavior when the correct
	        inputs are provided but is subject to input
	        injections
	\item $\dr{\beliefRegion}{\factRegion}$ the component
		never performs the expected behavior (e.g. physical
		damage)
\end{itemize}

\paragraph{Requirements as Facts}
During the specification phase, for any agent, channel, port, functional block
and architecture, there may exist a factorial requirement (Fact) predicating over them.
In other words, any requirement is defined as a Fact since they must be true in
any design or implementation. As we stated in Section~\ref{sec:theory} and depicted in
Figure~\ref{fig:soundness}, Facts are strategic rules that defines how
the system shall behave (by specification), while reality may be shown to
be insecure.

\begin{figure}[t]
	\centering
	\includegraphics[width=0.4\columnwidth]{sum.pdf}
	\caption{Example of a Functional Block}
	\label{fig:sum}
\end{figure}

As an example, consider a functional block that performs the summation of two
inputs, as in Figure~\ref{fig:sum}, defined by the requirement $r := b_3 = b_1
+ b_2$. The possible relations between the behavior of the functional block and
the requirements (i.e. $\Rcc{\beliefRegion}{\factRegion}$ is determined by
the relations between the I/O Beliefs $sum(b_1,b_2)=b_3$ and the requirement
$sum(b_1,b_2)=b_1+b_2$, as follows.
\begin{itemize}
	\item $\eq{\beliefRegion}{\factRegion}=\eq{\beliefRegion_3}{\beliefRegion_1+\beliefRegion_2}$, 
		where $\beliefRegion_3$ represents the Region of all the
		outputs of the functional block $sum$ while the Region
		$\beliefRegion_1+\beliefRegion_2$ represents the expected
		outputs of an ideal implementation of the requirement $r$.
		Therefore, the functional block correctly implements the
		requirements.
	\item
		$\dr{\beliefRegion}{\factRegion}=\dr{\beliefRegion_3}{\beliefRegion_1+\beliefRegion_2}$,
		the function block never implements the requirements.
	\item
		$\pp{\beliefRegion}{\factRegion}=\pp{\beliefRegion_3}{\beliefRegion_1+\beliefRegion_2}$, there exist some inputs for which
		the output is incorrect.
	\item $\ppi{\beliefRegion}{\factRegion}=\ppi{\beliefRegion_3}{\beliefRegion_1+\beliefRegion_2}$, there exists some outputs that are not the result of the summation of the inputs but given the correct inputs the function always outputs correctly.
	\item
		$\po{\beliefRegion}{\factRegion}=\po{\beliefRegion_3}{\beliefRegion_1+\beliefRegion_2}$,
		the component has a Byzantine behavior where occasionally
		outputs the expected output given the correct inputs. Not all
		the inputs are handled properly, nor all the expected outputs
		are always generated when correct inputs are given.
\end{itemize}

\subsection{Security and Insecurity of a System}
\begin{hypothesis}{\bf System Security Design --}\label{hyp:security}
	A system security design is given by a precise system 
	specification over the physical and functional architectures, that
	uniquely, i.e.  tested against the range of design possible in the
	$\abftheory$-theory, defines the design built on top of those
	requirements.
\end{hypothesis}

\begin{hypothesis}{\bf System Insecurity Design --}\label{hyp:insecurity}
	If, given a system specification as a collection of
	requirements on the system, there exist
	a non-unique design with respect to those requirements, the number of
	equivalent designs that fulfills the requirements quantitatively
	defines the magnitude of insecurity of a system design.
\end{hypothesis}

Based on these hypothesis, security ($S$) can be expressed as the following equation.
\begin{displaymath}
	S=1-(\prod_{a\in Ag} r^{\binom{n}{k}}-\delta_T)
\end{displaymath}
where $r^{\binom{n}{k}}$ represents the number of possible configurations (permutations) 
of an agent based on the number $r$ of relations (e.g. EQ, DR, PO) with arity $k$ between
$n$ different sets in the $\abftheory$ theory. $\delta_T$ represents the 
number of configurations which are not satisfiable with respect to the logical theory
defining the 
algebraic structure (topology) and constraints of the calculus (RCC). $Ag$ is the 
set of all the agents in the system under evaluation.
The constant 1 represents the nominal configuration of the system and then the
only secure configuration. The product represents all the unintended, insecure, configurations.

\subsection{Security Risk Assessment}\label{sec:secra}
In order to test our theory we implemented a tool-chain (open-source under the
AGPLv3 license and available at \autocite{cybersecurityGitHubRepo}) for the
identification of weaknesses and the precise calculation of potential insecure
configurations. The engineering of the $\abftheory$ is summarized in the UML
Class diagram in Figure~\ref{fig:secraclassdiagram} (in appendix).
